---
title: "Time Series Analysis"
author: "Hao Li"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Time Analysis sample code Rmd template


# Seasonality

Template for seasonality effect analysis with Holt-Winter model. With r


```{r forcastSetting}
nf=11#n of forcast
```
## Seasonal Plot

### Linear seasonal plot

```{r seasPlot,echo = FALSE}
n = length(UKgas)

m = matrix(as.matrix(UKgas),nrow = 4)
matplot(1:4,m,type = 'l',xlab = "Seasons", ylab =paste( "Seasonal value of ",deparse(substitute(UKgas))))

```


### Perspective Plot

```{r perspective}
persp(x=1:4,y=(1:(as.integer(n/4))),z = m,xlab = "Seasons", ylab ="Trend", zlab =paste( "Seasonal value of ",deparse(substitute(UKgas))) )

```


```{r perspPlot3D}
library(plot3D)

Mesh = mesh(1:4,(1:as.integer(n/4)))
Mesh

plot3D::surf3D(Mesh$x,Mesh$y,m,alpha = .5,main = "Seasonal 3D View")


vx = as.vector(Mesh$x) ; vy = as.vector(Mesh$y); vz = as.vector(m)

#yMesh<-zMesh<-matrix(NA,nrow)



t = 1:length(UKgas)
r = as.vector(UKgas)
z = r*cos(t)
x = r*sin(t)
y = t
lines3D(x,y,z,main = "Cylindrical view of seasonal data")


```

### Polar Seasonal Plot


```{r polarSeasonal}

library(ggplot2)
#install.packages("forecast")
library(forecast)
ggseasonplot(UKgas,polar = F)
ggseasonplot(UKgas,polar = T)

```


## Decomposition

```{r, echo = FALSE}

dts_add =  decompose(UKgas,type ="additive")
plot(dts_add);grid()

plot(dts_add$figure,ylab = "Seasonality index",main  = "Seasonality of the additive model", type = 'b',pty = 5);grid()
dts_mul = decompose(UKgas,type = "multiplicative")
plot(dts_add$figure,ylab = "Seasonality index",main  = "Seasonality of the multiplicative model", type = 'b');grid()
plot(dts_mul);grid()


```


## 2 Holt-Winter model (Ets model)

```
```


### Addictive trend, addictive seasonality

### Addictive trend, multiplicative seasonality

### Multiplicative trend, addictive seasonality

### Multiplicative trend, multiplicative seasonality

# Lag

```{r}

n=length(UKgas)
t=seq(1,n,1)
#2. Data manipulation
#repeat for Autoregression involving 1:3 previous terms
#(to avoid copying and pasting lag1=...,fit1...)

UKgasdf = data.frame(t=t,UKgas=UKgas)
UKgasdf = within(UKgasdf,{
  for(i in 1:3) assign(paste0('lag',i),c(rep(NA,i),UKgas[1:(n-i)]))#assign lag1:lag3
  remove(i)
})
require(ggplot2)
ggplot(data = UKgasdf,aes(x = lag1,y = UKgas)) + geom_point(alpha = .4) + geom_rug(alpha = .4)
ggplot(data = UKgasdf,aes(x = lag2,y = UKgas)) + geom_point(alpha = .4) + geom_rug(alpha = .4)
ggplot(data = UKgasdf,aes(x = lag3,y = UKgas)) + geom_point(alpha = .4) + geom_rug(alpha = .4)

require(plot3D)
plot3D::scatter3D(UKgasdf$lag2,UKgasdf$lag1,UKgasdf$UKgas)
```

```{r}
fit1 = lm(UKgas~t+lag1,data=UKgasdf)
fit2 = lm(UKgas~t+lag1+lag2,data = UKgasdf)
fit3 = lm(UKgas~t+lag1+lag2+lag3,data = UKgasdf)
require(visreg)
print("1 lag model")
summary(fit1)
visreg::visreg(fit1)
visreg2d(fit2,xvar = 't',yvar = 'lag1',plot.type = "persp")
print("2 lag model")
summary(fit2)
visreg2d(fit2,xvar = 'lag1',yvar = 'lag2',plot.type = "persp")
#visreg2d
print("3 lag model")
summary(fit3)
visreg2d(fit3,xvar = 'lag1',yvar = 'lag2',plot.type = "persp")
visreg2d(fit3,xvar = 'lag2',yvar = 'lag3',plot.type = "persp")
visreg2d(fit3,xvar = 'lag1',yvar = 'lag3',plot.type = "persp")
```

### Residuals

```{r,echo = FALSE}
UKgasdf
UKgasdf = within(UKgasdf,{
  for(i in 1:3) assign(paste0('fit',i),c(rep(NA,i),predict.lm(eval(parse(text = paste0('fit',i))))))#assign lag1:lag3
  remove(i)
})
print(head(UKgasdf))

```


```{r}
UKgasdf =within(UKgasdf,{
  for(i in 1:3) assign(paste0('res',i),UKgas - get(paste0('fit',i)))#assign lag1:lag3
  remove(i)
})
#UKgasdf
head(UKgasdf)
layout(1)
UKgaspred = c('UKgas','fit1','fit2','fit3')
matplot(x=UKgasdf[,'t'],y=UKgasdf[,UKgaspred],type = c('h','l','l','l'),lty = 1:4,col =1:4,
        xlab = 't',ylab = 'value')
grid()
legend(0,9,legend = UKgaspred,lwd=2,col = 1:4,lty = 1:4)

#layout(1:3)
for(i in 1:3) hist(UKgasdf[,as.character(paste0('res',i))],xlab =  paste0('residual',i),main = paste0('residual',i))
layout(1)
sd_error=numeric(3)
for(i in 1:3) sd_error[i] = sd(UKgasdf[,as.character(paste0('res',i))],na.rm =T)
plot(c(1:3), sd_error,
     xlab = 'No. of previous terms(complexity)',ylab = 'SD of residual',type = 'b')
#plot shows how residuals get lowered at the expense of greater complexity
print("Does residuals get lowered at the expense of greater complexity?")

```



```{r}
#uncomment the code below to avoid copying and pasting
#!!!Note that this is usually not encouraged because it is unintuitive
#(to avoid copying and pasting lag1=...,fit1...)
#for(i in 1:3){
#  indp = paste0('lag',1:i)
#  sapply(indp,paste,sep = "+")
#  assign(paste0('fit',i),lm(as.formula(paste('UKgas ~ t +',indp))))
#}
#  for(i in 1:3) print(summary(eval(parse(text = paste0('fit',i)))))


# plot autocovariance function for these residuals
#require(graphics)
#layout(1:3)
for(i in 1:3) acf(UKgasdf[(i+1):nrow(UKgasdf),paste0('res',i)])
#acf(e1)


# forecast UKgas for May 1990
for(i in 1:3) {
  assign(paste0('coef',i),coef(get(paste0('fit',i))))
 print(get(paste0('coef',i)))
 }

t2=seq(1,n+nf,1)

for(i in 1:3){
  #for every model
  UKgasf = c(as.vector(UKgas),numeric(nf))
  #for(k in 1:i){
   #   #for every lag
    #  coeflag = c(coeflag,get(paste0('coef',k))[paste0('lag',k)])
    #}
  for(j in n+1:nf){
    #for every ts term
    

    #UKgasflag = NULL
    #for(k in 1:i){
      #for every lag
    UKgasflag= UKgasf[(j-i):(j-1)]
    UKgasf[j] = sum(get(paste0('coef',i))*c(1,t2[j],UKgasflag))
  }
  assign(paste0('UKgasf',i),UKgasf)
  #fit2 = numeric(length(UKgasf))
  #for(j in n+i:nf){
  #  fit2[j] = sum(get(paste0('coef',i))*c(1,t2[j],UKgasf[(j-i+1):(j)]))
  #}
  assign(paste0('fit2_',i),fit2)
}
#layout(1:3)
for(i in 1:3){
  plot(t2,get(paste0('UKgasf',i)),ylab=paste0('fit2_',i), type ='b')
  points(t,UKgas,col = 'red')
  abline(v = n)
}

layout(1)
UKgaspred = c('UKgas','fit1','fit2','fit3')
matplot(x=UKgasdf[,'t'],y=UKgasdf[,UKgaspred],type = c('h','l','l','l'),lty = 1:4,col =1:4,
        xlab = 't (month)',ylab = 'value',
        xlim = c(0,n+nf),
        ylim = c(min(UKgas) - 1.5*sd(UKgas),
                 (max(UKgas)+1.5*sd(UKgas))))
grid()
legend(0,(min(UKgas)+max(UKgas))/2,bg = NULL,legend = UKgaspred,lwd=2,col = 1:4,lty = 1:4)
for(i in 1:3){
  lines(t2,get(paste0('UKgasf',i)),ylab=paste0('fit2_',i), type ='l',col = i+1)
  #points(t,UKgas,col = 'red')
  abline(v = n,col = 'purple',lty = 3)
}
lines(t,UKgas,col = 'grey')


for(i in 1:3) {
  print(i)
  print(get(paste0('UKgasf',i)))
}

```


# Filter

## Linear


```{r, predictPolynomial}
predictPolynomial = function(model,t){
  makeTpower = function(t,p){
    result = c()
    for(i in 0:p) result = c(result,t^i)
    return(result)
  }
  coef_model = coefficients(model)
  n_order = length(coef_model)
  result = c()
  for(i in seq_along(t)) result = c(result,sum(makeTpower(t,n_order)*coef_model))
  result
}

```



```{r fitLog}
fitLin = lm(UKgas ~ t)
```
### Linear model and residuals
```{r}
visreg::visreg(fitLin)
resLin = residuals(fitLin)
plot(resLin,xlab = "t", ylab = "Residual of Linear Model",type = 'h');grid()

```

## Generalized_linear - quadratic

```{r}
t_sqr=t^2
fitQuad = lm(UKgas~ t+t_sqr)
summary(fitQuad)
visreg::visreg(fitQuad)
plot3D::scatter3D(t,t_sqr,UKgas)
visreg2d(fitQuad,'t','t_sqr',plot.type = "persp")
resLin = residuals(fitQuad)
plot3D::scatter3D(t,t_sqr,resLin)
```
### generalized_linear - cubed

```{r}
t_cubed = t^3
fit_cubed = lm(UKgas~ t+t_sqr+t_cubed)
summary(fit_cubed)
visreg::visreg(fit_cubed)
coefficients(fit_cubed)


```


## Exponential

## Logarithm
```{r fitLog}
fitLog = lm(UKgas ~ (t))
```


# Periodicity

```{r fft}
plot(fft(UKgas))
require(psd)
pspectrum(UKgas,plot = TRUE)
```

# Non-linearity

```{r maxLyapunowExponent}

```

## Fractal properties

```{r fractal}
#install.packages("fractal")
#install.packages("tseriesChaos")
#install.packages("nonlinearTseries")
#install.packages("timedelay")
library(timedelay)
library(nonlinearTseries)
library(tseriesChaos)
library(fractal)

determinism(UKgas)
spaceTime(UKgas)

dim_emb = estimateEmbeddingDim(UKgas)
print(paste("Estimated embedding dimension",dim_emb))

lag_est = timeLag(UKgas)
print(paste("Estimation of time lag:",lag_est))
#lyap_k(UKgas,m =dim_emb ,d =timeLag(UKgas),ref = ceiling(length(UKgas)/3))

n.reference = min(as.integer(round(length(UKgas)/20)), 100)
ifelse(n.reference>=10,print("Valid"),{message("Length may be insufficient, resetting n.reference to 10");n.reference = 10})
#lya_est = lyapunov(UKgas, tlag=lag_est, dimension=dim_emb, local.dimension=1, n.reference = n.reference)
#print(lya_est)


```



```{r fractalDimesion}

#corrDim(UKgas,min.radius = 1/50*sd(UKgas),max.radius = 1/25*sd(UKgas))
#infoDim(UKgas)

```

```{r entropy}
#install.packages("pracma")
library(pracma)

r_vec = seq(from = .01 ,to =1, by=.01)


entropy_est = numeric(length(r_vec))
for(i in seq_along(r_vec)) entropy_est[i]<-approx_entropy(UKgas, edim = dim_emb, r = r_vec[i]*sd(UKgas), elag = lag_est)
plot(r_vec,entropy_est,xlab = "r / Standard Deviation",ylab = "Estimated Entropy",type = 'b');grid()
abline(v = r_vec[which.max(entropy_est)])
max(entropy_est)
```



